### 2025-03-14 paper/sections/related.qmd
Parse key comments from second/third document
TODO Patrick: remaining comments, namely (1) better labels for Fig. 1? (2) reorganizing Table 1? (3) potentially some extra explanation in results?
Back to (exactly) 16 pages!
TODO: Complete read-through

This interpretation of the link between explainability through counterfactuals on one side and robustness to adversarial examples on the other is backed by empirical evidence. @sauer2021counterfactual demonstrate that using counterfactual images during classifier training improves model robustness. Similarly, @abbasnejad2020counterfactual argue that counterfactuals represent potentially useful training data in machine learning, especially in supervised settings where inputs may be reasonably mapped to multiple outputs. They, too, demonstrate that augmenting the training data of image classifiers can improve generalization. Finally, @teney2020learning propose an approach using counterfactuals in training that does not rely on data augmentation: they argue that counterfactual pairs typically already exist in training datasets. Specifically, their approach relies on identifying similar input samples with different annotations and ensuring that the gradient of the classifier aligns with the vector between such pairs of counterfactual inputs using the cosine distance as the loss function.

In the natural language processing (NLP) domain, CEs have also been used to improve models through data augmentation. @wu2021polyjuice propose *Polyjuice*, a general-purpose counterfactual generator for language models. The authors empirically demonstrate that the augmentation of training data with their method improves robustness in a number of NLP tasks. @balashankar2023improving similarly use *Polyjuice* to augment NLP datasets through diverse counterfactuals and show that classifier robustness improves by up to 20%. Finally, @luu2023counterfactual introduce Counterfactual Adversarial Training (CAT), which also aims at improving generalization and robustness of language models through a three-step procedure: the authors identify training samples that are subject to high predictive uncertainty, generate CEs for those samples, and then fine-tune the language model on a dataset augmented with the CEs.

>>>>

This interpretation of the link between explainability through counterfactuals on one side and robustness to adversarial examples on the other is backed by empirical evidence. @sauer2021counterfactual demonstrates that using counterfactual images during classifier training improves model robustness. Similarly, @abbasnejad2020counterfactual argue that counterfactuals represent potentially useful training data in machine learning, especially in supervised settings where inputs may be reasonably mapped to multiple outputs. They, too, demonstrate that augmenting the training data of image classifiers can improve generalization. Finally, @teney2020learning propose an approach using counterfactuals in training that does not rely on data augmentation: they argue that counterfactual pairs typically already exist in training datasets. Specifically, their approach relies on identifying similar input samples with different annotations and ensuring that the gradient of the classifier aligns with the vector between such pairs of counterfactual inputs using the cosine distance as the loss function.

In the natural language processing (NLP) domain, CEs have also been used to improve models through data augmentation. @wu2021polyjuice2 propose *Polyjuice*, a general-purpose counterfactual generator for language models. The authors empirically demonstrate that the augmentation of training data with their method improves robustness in a number of NLP tasks. @balashankar2023improving similarly use *Polyjuice* to augment NLP datasets through diverse counterfactuals and show that classifier robustness improves by up to 20%. Finally, @luu2023counterfactual introduce Counterfactual Adversarial Training (CAT), which also aims at improving generalization and robustness of language models through a three-step procedure: the authors identify training samples that are subject to high predictive uncertainty, generate CEs for those samples, and then fine-tune the language model on a dataset augmented with the CEs.
